Impact of Federal Party Leadership Changes on Donation Patterns

### 2024-12-04 paper/paper.qmd
Update model tests

### Assumptions and Limitations
The models assume:

Linearity: The relationship between predictors and the log of total donations is linear.
Normality of Errors: The residuals are normally distributed.
Homoscedasticity: Constant variance of residuals.
No Multicollinearity: Predictors are not highly correlated.
Independence of Observations: Each observation is independent.
Potential limitations include unobserved variables affecting donation amounts and the assumption that past trends predict future behavior.

### Model Validation 
Model validation was conducted to assess the predictive accuracy and generalizability of the regression models. This involved evaluating key metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square Error (RMSE) to determine how well the models explain the variance in Log_Total_Donations and their predictive performance.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-model-summary
#| tbl-cap: "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models"
#| echo: false
#| message: false

library(broom)
library(knitr)
library(kableExtra)

ontario_data <- analysis_data %>% filter(region == "Ontario")

# Aggregate total donations per party per year in Ontario
total_donations_ontario <- ontario_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Run Regression for Ontario with Controls

total_donations_ontario <- total_donations_ontario %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)

model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)

federal_data <- analysis_data %>% filter(region == "Federal")

# Aggregate total donations per party per year at Federal level
total_donations_federal <- federal_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Run Regression for Federal with Controls
# For Federal
total_donations_federal <- total_donations_federal %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))

model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)

cl_data <- analysis_data %>%
  filter(political_party %in% c(
    "Conservative Party of Canada",
    "Liberal Party of Canada",
    "Progressive Conservative Party of Ontario",
    "Liberal Party of Ontario"
  ))

# Create variables
cl_data <- cl_data %>%
  mutate(
    In_Power = recipient_in_power,  # Already binary (1/0)
    Party = ifelse(
      political_party %in% c("Conservative Party of Canada", "Progressive Conservative Party of Ontario"),
      1,  # Conservative
      0   # Liberal
    )
  )

# Aggregate total donations per party per year
total_donations_cl <- cl_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(In_Power),
    Party = first(Party),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Log-transform the dependent variable to stabilize variance
total_donations_cl <- total_donations_cl %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))

# Run the regression with interaction term
model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)

# Tidy model summaries
tidy_ontario <- tidy(model_ontario) %>%
  mutate(Model = "Ontario") %>%
  select(Model, term, estimate, std.error, p.value)

tidy_federal <- tidy(model_federal) %>%
  mutate(Model = "Federal") %>%
  select(Model, term, estimate, std.error, p.value)

tidy_cl <- tidy(model_cl) %>%
  mutate(Model = "Conservative/Liberal") %>%
  select(Model, term, estimate, std.error, p.value)

# Combine tidy summaries
tidy_all <- bind_rows(tidy_ontario, tidy_federal, tidy_cl)

# Round numeric values to 3 decimal places
tidy_all <- tidy_all %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    p.value = round(p.value, 3)
  )

# Create a formatted table
kable(tidy_all, 
      caption = "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models",
      col.names = c("Model", "Term", "Estimate", "Std. Error", "p-value"),
      align = "lcccc") %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  add_header_above(c(" " = 1, "Regression Coefficients" = 4)) %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(5, color = "red")

>>>>

### Assumptions and Limitations
The models assume a linear relationship between the predictors and the logarithm of total donations, ensuring that the effects of the predictors are additive and proportional. Additionally, they assume that the residuals are normally distributed (normality of errors) and exhibit constant variance (homoscedasticity). The models also rely on the absence of multicollinearity, meaning that the predictors are not highly correlated, and they assume independence of observations, where each data point is independent of the others. Potential limitations of these models include the possibility of unobserved variables influencing donation amounts and the inherent assumption that past trends can reliably predict future behavior.

### Model Validation ******
Model validation was conducted to assess the predictive accuracy and generalizability of the regression models. This involved evaluating key metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square Error (RMSE) to determine how well the models explain the variance in Log_Total_Donations and their predictive performance.

@@
